Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2016-03-02 draft.tex
remove lowercase state variables throughout body of paper

The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors. That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.\footnote{In the case of $w_s^t$ and $w_g^t$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables. However, positive investments in some of the state variables can be ruled out by preference assumptions.

Player $c$ dislikes war and so will never invest in $w_s$ or $w_g$ since they increase the likelihood that one of the inside actors chooses to fight. It would also not want to make the government lose and so won't invest in $l_g$. This leaves three state variables in which player $c$ might invest: $q_s$, $q_g$, $l_s$.

Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $w_g$ or $l_s$. The only reason the patron would invest in $q_g$ is to counter an investment by player $c$ in the government's payoffs from war; since player $c$ will not make such an investment, we can also rule out investments by the patron in $q_g$. This leaves four state variables in which the patron may invest: $q_s$, $l_g$ and $w_s$.

@@

\begin{enumerate}
	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),0\right\}$ to augment $q_s^t$. Player $c$ invests $R_c^t =0$ in $l_s^t$.

If $R_p^t < \frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right)$ player $c$ invests $R_c^t = l_s^t - \left(q_s^t + R_p^t \right) + \ve$, for $\ve$ small, in $l_s^t$. Cede becomes the unique best response for the secessionists.

	\item As long as $R_p^t \leq \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = R_p^t - \left( Q_g^{t-1} - L_g^{t-1}\right)$ in $q_g^t$. Note that the outcome is that investments in $l_g^t$ and $q_g^t$ are zero.

If $R_p^t > \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = 0$. Cede becomes the unique best response for the government.

	\item As long as $R_p^t \leq \frac{1}{p_s} \left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_s + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)$ in $q_s^t$. Similar to case two, the outcome is that investments in $w_s^t$ and $q_s^t$ are zero.

If $R_p^t > \frac{1}{p_s}\left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$. Fight becomes the unique best response for the secessionists.
\end{enumerate}
\end{definition}

@@

Recall that either outside actor could invest toward increasing any of the six state variables in a period: $q_s^t,q_g^t,w_s^t,w_g^t,l_s^t,$ and $l_g^t$. However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2. The remainder of the proof addresses each set of matched investment decisions in turn.

@@

Consider period 1. When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1 \ (= q_s^1 + R_s^1)$ and $l_s^1$ so that $c$ will not invest. That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(q_s^1 - l_s^1)$ as long as $p$ can afford to make this investment. Thus we need Restriction (5): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $t > 1$, there are two cases to consider. Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$. Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$. However, because in each period $q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo (i.e. $q_s^{t+1} = Q_s^t - \mu -\sigma$ where $\sigma$ is the reduction due to sanctions).  In order for sanctions to lead to ceding by the secessionists, the following are required:

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e. $q_s^{t+1} = Q_s^t - \mu -\sigma$ \emph{and} $p_s\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$). The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
\end{proposition}

>>>>

The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors. That is, $s = \left(Q_s,Q_g,L_s,L_g,W_s,W_g \right)$.\footnote{In the case of $W_s^t$ and $W_g^t$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables. However, positive investments in some of the state variables can be ruled out by preference assumptions.

Player $c$ dislikes war and so will never invest in either state variable associated with winning since they increase the likelihood that one of the inside actors chooses to fight. It would also not want to make the government lose and so won't invest to augment $L_g$. This leaves three state variables in which player $c$ might invest: $Q_s$, $Q_g$, $L_s$.

Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $W_g$ or $L_s$. The only reason the patron would invest in $Q_g$ is to counter an investment by player $c$ in the government's payoffs from war; since player $c$ will not make such an investment, we can also rule out investments by the patron in $Q_g$. This leaves four state variables in which the patron may invest: $Q_s$, $L_g$ and $W_s$.

@@

\begin{enumerate}
	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),0\right\}$ to augment $Q_s^{t-1}$. Player $c$ invests $R_c^t =0$ to augment $L_s^{t-1}$.

If $R_p^t < \frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right)$ player $c$ invests $R_c^t = L_s^{t-1} - \left(Q_s^{t-1} - \mu + R_p^t \right) + \ve$, for $\ve$ small, to augment $L_s^{t-1}$. Cede becomes the unique best response for the secessionists.

	\item As long as $R_p^t \leq \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $L_g^{t-1}$, player $c$ invests $R_c^t = R_p^t - \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $Q_g^{t-1}$. Note that the outcome is that investments to augment $L_g^{t-1}$ and $Q_g^{t-1}$ are zero.

If $R_p^t > \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $L_g^{t-1}$, player $c$ invests $R_c^t = 0$. Cede becomes the unique best response for the government.

	\item As long as $R_p^t \leq \frac{1}{p_s} \left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right]$ to augment $W_s^{t-1}$, player $c$ invests $R_c^t = R_p^tp_s + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)$ to augment $Q_s^{t-1}$. Similar to case two, the outcome is that investments to augment $W_s^{t-1}$ and $Q_s^{t-1}$ are zero.

If $R_p^t > \frac{1}{p_s}\left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right] $ to augment $W_s^{t-1}$, player $c$ will choose $R_c^t=0$. Fight becomes the unique best response for the secessionists.
\end{enumerate}
\end{definition}

@@

Recall that either outside actor could invest toward increasing any of the six state variables in a period. However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2. The remainder of the proof addresses each set of matched investment decisions in turn.

@@

Consider period 1. When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1$ and $L_s^1$ so that $c$ will not invest. That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(Q_s^0 - \mu - L_s^0)$ as long as $p$ can afford to make this investment. Thus we need Restriction (5): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $t > 1$, there are two cases to consider. Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$. Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$. However, because in each period $Q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo (i.e. $Q_s^t = Q_s^{t-1} - \mu -\sigma +R_p^t - R_c^t$ where $\sigma$ is the reduction due to sanctions).  In order for sanctions to lead to ceding by the secessionists, the following are required:

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e. $Q_s^t = Q_s^{t-1} - \mu -\sigma +R_p^t - R_c^t$ \emph{and} $p_s\left(\sigma\right)$ $\left(p_g\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$). The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
\end{proposition}
